Quasi-Monte Carlo Methods: What, Why, and How?

### 2025-01-08 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

\begin{abstract}
Many problems in  quantitative finance, uncertainty quantification, and other areas have answers of the form $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation. The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\boldsymbol{X}_i)$ for a well chosen sequence of nodes, $\{\bsX_0, \bsX_1, \ldots\}$.  Computing $\mu$ is equivalent to computing a $d$-dimensional integral.

@@

In mathematical terms, the outcome is $Y := f(\boldsymbol{X})$, where $\boldsymbol{X}$ is a vector random variable of inputs determining the outcome.  The values of $f$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.  The scenarios are $\bsX_0, \bsX_1, \ldots$, which give rise to outcomes $Y_0 = f(\bsX_0), Y_1 = f(\bsX_1), \ldots$.  The $Y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest. These scenarios are also commonly called instances of $\bsX$ or nodes.  We use these terms interchangeably.

Simple or independent and identically distributed (IID)  Monte Carlo chooses the sequence of nodes $\{\bsX_0, \bsX_1, \ldots\}$ to be IID.  Quasi-Monte Carlo (qMC) methods choose the nodes to be \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.  The difference between these two distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy sequences.

This tutorial describes what qMC is, why we would want to use qMC, and how qMC can be implemented well.  The next session provides an illustrative example of the advantages of qMC.  This is followed by a description of how low discrepancy sequences are constructed and various measures of discrepancy.  We then explain how randomization can improve qMC, how to decide what sample size, $n$, is sufficient, and how to rewrite the problem of interest in a qMC-friendly way.

@@

The aim is to approximate $\mu$ by a sample mean,
\begin{equation} \label{eq:sample_mean}
	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\boldsymbol{X}_i),
\end{equation}
but to do so requires some further preparation.  This integral above, $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
\begin{equation}\label{eq:keisterB}
	\mu = \int_{\mathbb{R}^d} \underbrace{\pi^{d/2} \cos(\lVert \bst \rVert)}_{=:g(\bst)} \cdot \underbrace{\frac{\exp(-\lVert \bst \rVert^2)}{\pi^{d/2}}}_{\text{density of } \calN(\bszero,\mathsf{I}/2)} \, \rmd \bst = \bbE(Y) = \bbE[g(\bsT)].
\end{equation}

Since the low discrepancy sequences that underly qMC are defined to approximate the standard uniform distribution, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.  This reimagines the integral $\mu$ as the expectation of a function, $f$, of a standard uniform random variable:
\begin{multline}\label{eq:keisterC}
	\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d)\bigr) \bigr\rVert/\sqrt{2}  \Bigr)}_{=:f(\bsx)} \, \rmd \bsx \\
	= \bbE[Y]
	= \bbE[f(\bsX)], \qquad \bsX \sim \calU[0,1]^d.
\end{multline}

Consider the specific case of $d=6$ for which $\mu \approx -2.327303729298$.  We approximate $\mu$ by the sample mean defined in \eqref{eq:sample_mean} for three different kinds of nodes, $\{\bsX_i\}_{i=0}^{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}. The three kind of nodes are
\begin{enumerate}
	\renewcommand{\labelenumi}{\roman{enumi}.}
	\item Cartesian grids, $\{1/(2m), \ldots, (2m-1)/(2m) \}^6$, for $m = 2, 3, \ldots$ (blue),
	\item IID simple random  with arbitrary $n$ (orange), and
	\item Randomized low discrepancy (Sobol') sequences with $n = 1, 2, \ldots, 2^m, \ldots $ (green).
\end{enumerate}

\begin{figure}
	\centering
	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsX_0, \bsX_1, \ldots$.  Grids have the largest error and low discrepancy (qMC) nodes have the smallest error. \label{fig:keister-err}}
\end{figure}

Note the following from this example:
\begin{itemize}
	\item This integral is not particularly easy.  Even the best choice of nodes requires at least $n=100$ to get a relative error below $10\%$.

	\item \emph{Grid nodes.} Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.  Figure \ref{} shows a grid with $n = 2$, $3$, and $6$For this modest $d$, the possible sample sizes, $n$, are quite sparse, as shown in Figure \ref{fig:keister-err}.

@@

	\item \emph{IID nodes.}  Simple MC (orange) is a substantial improvement over grid nodes. As a reminder, for IID nodes the root mean squared error is
	\begin{equation}\label{eq:IIDerror}
		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsX))}{n}} = \frac{\mathrm{Std}(f(\bsX))}{n^{1/2}},
	\end{equation}
	where $\var$ denotes the variance and $\mathrm{Std}$ the standard deviation.  This $\mathcal{O}(n^{-1/2})$ decay is observed in Figure \ref{fig:keister-err}. For simple MC the sample size, $n$, can be any positive integer without affecting the rate of decay of the error.

	\item \emph{Low discrepancy nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.  Typically, low discrepancy sequences have preferred sample sizes.  For this Sobol' sequence, $n$ equal to a positive integer power of $2$ tends to give better accuracy than arbitrary $n$.

@@

\section{Low Discrepancy Sequence Constructions} \label{sec:construct}
\section{Discrepancy} \label{sec:discrepancy}
\section{Randomization} \label{sec:random}
\section{Stopping Criteria} \label{sec:stop}
\section{Reformulating Our Problem} \label{sec:reformulate}
\section{Ongoing Research} \label{sec:onging}
\section{Conclusion} \label{sec:conclusion}

@@

\end{document}
%
In this section, we will consider the case when the Hamiltonian $H(x)$
is autonomous. For the sake of simplicity, we shall also assume that it
is $C^{1}$.

We shall first consider the question of nontriviality, within the
general framework of
$\left(A_{\infty},B_{\infty}\right)$-subquadratic Hamiltonians. In
the second subsection, we shall look into the special case when $H$ is
$\left(0,b_{\infty}\right)$-subquadratic,
and we shall try to derive additional information.
%
\subsubsection{The General Case: Nontriviality.}
%
We assume that $H$ is
$\left(A_{\infty},B_{\infty}\right)$-sub\-qua\-dra\-tic at infinity,
for some constant symmetric matrices $A_{\infty}$ and $B_{\infty}$,
with $B_{\infty}-A_{\infty}$ positive definite. Set:
\begin{eqnarray}
\gamma :&=&{\rm smallest\ eigenvalue\ of}\ \ B_{\infty} - A_{\infty} \\
  \lambda : &=& {\rm largest\ negative\ eigenvalue\ of}\ \
  J \frac{d}{dt} +A_{\infty}\ .
\end{eqnarray}

Theorem~\ref{ghou:pre} tells us that if $\lambda +\gamma < 0$, the
boundary-value problem:
\begin{equation}
\begin{array}{rcl}
  \dot{x}&=&JH' (x)\\
  x(0)&=&x (T)
\end{array}
\end{equation}
has at least one solution
$\overline{x}$, which is found by minimizing the dual
action functional:
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
\end{equation}
on the range of $\Lambda$, which is a subspace $R (\Lambda)_{L}^{2}$
with finite codimension. Here
\begin{equation}
  N(x) := H(x) - \frac{1}{2} \left(A_{\infty} x,x\right)
\end{equation}
is a convex function, and
\begin{equation}
  N(x) \le \frac{1}{2}
  \left(\left(B_{\infty} - A_{\infty}\right) x,x\right)
  + c\ \ \ \forall x\ .
\end{equation}

%
\begin{proposition}
Assume $H'(0)=0$ and $ H(0)=0$. Set:
\begin{equation}
  \delta := \liminf_{x\to 0} 2 N (x) \left\|x\right\|^{-2}\ .
  \label{eq:one}
\end{equation}

If $\gamma < - \lambda < \delta$,
the solution $\overline{u}$ is non-zero:
\begin{equation}
  \overline{x} (t) \ne 0\ \ \ \forall t\ .
\end{equation}
\end{proposition}
%
\begin{proof}
Condition (\ref{eq:one}) means that, for every
$\delta ' > \delta$, there is some $\varepsilon > 0$ such that
\begin{equation}
  \left\|x\right\| \le \varepsilon \Rightarrow N (x) \le
  \frac{\delta '}{2} \left\|x\right\|^{2}\ .
\end{equation}

It is an exercise in convex analysis, into which we shall not go, to
show that this implies that there is an $\eta > 0$ such that
\begin{equation}
  f\left\|x\right\| \le \eta
  \Rightarrow N^{\ast} (y) \le \frac{1}{2\delta '}
  \left\|y\right\|^{2}\ .
  \label{eq:two}
\end{equation}

\begin{figure}
\vspace{2.5cm}
%%%You can un-comment the \includegraphics{} line below and comment the one above
%%%if you have copied in your directory the file figure.eps
%%%from Springer zip file. As mentioned in the preamble, in your manuscript you are encouraged to put
%%%figures and images in a sub-directory which can be provided in the preamble via the
%%%command \graphicspath{{<path 1>}{<path 2>}}
%\includegraphics{figure.eps}
\caption{This is the caption of the figure displaying a white eagle and
a white horse on a snow field}
\end{figure}

Since $u_{1}$ is a smooth function, we will have
$\left\|hu_{1}\right\|_\infty \le \eta$
for $h$ small enough, and inequality (\ref{eq:two}) will hold,
yielding thereby:
\begin{equation}
  \psi (hu_{1}) \le \frac{h^{2}}{2}
  \frac{1}{\lambda} \left\|u_{1} \right\|_{2}^{2} + \frac{h^{2}}{2}
  \frac{1}{\delta '} \left\|u_{1}\right\|^{2}\ .
\end{equation}

If we choose $\delta '$ close enough to $\delta$, the quantity
$\left(\frac{1}{\lambda} + \frac{1}{\delta '}\right)$
will be negative, and we end up with
\begin{equation}
  \psi (hu_{1}) < 0\ \ \ \ \ {\rm for}\ \ h\ne 0\ \ {\rm small}\ .
\end{equation}

On the other hand, we check directly that $\psi (0) = 0$. This shows
that 0 cannot be a minimizer of $\psi$, not even a local one.
So $\overline{u} \ne 0$ and
$\overline{u} \ne \Lambda_{o}^{-1} (0) = 0$. \qed
\end{proof}
%
\begin{corollary}
Assume $H$ is $C^{2}$ and
$\left(a_{\infty},b_{\infty}\right)$-subquadratic at infinity. Let
$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  be the
equilibria, that is, the solutions of $H' (\xi ) = 0$.
Denote by $\omega_{k}$
the smallest eigenvalue of $H'' \left(\xi_{k}\right)$, and set:
\begin{equation}
  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
\end{equation}
If:
\begin{equation}
  \frac{T}{2\pi} b_{\infty} <
  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
  \frac{T}{2\pi}\omega
  \label{eq:three}
\end{equation}
then minimization of $\psi$ yields a non-constant $T$-periodic solution
$\overline{x}$.
\end{corollary}
%

We recall once more that by the integer part $E [\alpha ]$ of
$\alpha \in \bbbr$, we mean the $a\in \bbbz$
such that $a< \alpha \le a+1$. For instance,
if we take $a_{\infty} = 0$, Corollary 2 tells
us that $\overline{x}$ exists and is
non-constant provided that:

\begin{equation}
  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
\end{equation}
or
\begin{equation}
  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
  \label{eq:four}
\end{equation}

%
\begin{proof}
The spectrum of $\Lambda$ is $\frac{2\pi}{T} \bbbz +a_{\infty}$. The
largest negative eigenvalue $\lambda$ is given by
$\frac{2\pi}{T}k_{o} +a_{\infty}$,
where
\begin{equation}
  \frac{2\pi}{T}k_{o} + a_{\infty} < 0
  \le \frac{2\pi}{T} (k_{o} +1) + a_{\infty}\ .
\end{equation}
Hence:
\begin{equation}
  k_{o} = E \left[- \frac{T}{2\pi} a_{\infty}\right] \ .
\end{equation}

The condition $\gamma < -\lambda < \delta$ now becomes:
\begin{equation}
  b_{\infty} - a_{\infty} <
  - \frac{2\pi}{T} k_{o} -a_{\infty} < \omega -a_{\infty}
\end{equation}
which is precisely condition (\ref{eq:three}).\qed
\end{proof}
%

\begin{lemma}
Assume that $H$ is $C^{2}$ on $\bbbr^{2n} \setminus \{ 0\}$ and
that $H'' (x)$ is non-de\-gen\-er\-ate for any $x\ne 0$. Then any local
minimizer $\widetilde{x}$ of $\psi$ has minimal period $T$.
\end{lemma}
%
\begin{proof}
We know that $\widetilde{x}$, or
$\widetilde{x} + \xi$ for some constant $\xi
\in \bbbr^{2n}$, is a $T$-periodic solution of the Hamiltonian system:
\begin{equation}
  \dot{x} = JH' (x)\ .
\end{equation}

There is no loss of generality in taking $\xi = 0$. So
$\psi (x) \ge \psi (\widetilde{x} )$
for all $\widetilde{x}$ in some neighbourhood of $x$ in
$W^{1,2} \left(\bbbr / T\bbbz ; \bbbr^{2n}\right)$.

But this index is precisely the index
$i_{T} (\widetilde{x} )$ of the $T$-periodic
solution $\widetilde{x}$ over the interval
$(0,T)$, as defined in Sect.~2.6. So
\begin{equation}
  i_{T} (\widetilde{x} ) = 0\ .
  \label{eq:five}
\end{equation}

Now if $\widetilde{x}$ has a lower period, $T/k$ say,
we would have, by Corollary 31:
\begin{equation}
  i_{T} (\widetilde{x} ) =
  i_{kT/k}(\widetilde{x} ) \ge
  ki_{T/k} (\widetilde{x} ) + k-1 \ge k-1 \ge 1\ .
\end{equation}

This would contradict (\ref{eq:five}), and thus cannot happen.\qed
\end{proof}
%
\paragraph{Notes and Comments.}
The results in this section are a
refined version of \cite{smit:wat};
the minimality result of Proposition
14 was the first of its kind.

To understand the nontriviality conditions, such as the one in formula
(\ref{eq:four}), one may think of a one-parameter family
$x_{T}$, $T\in \left(2\pi\omega^{-1}, 2\pi b_{\infty}^{-1}\right)$
of periodic solutions, $x_{T} (0) = x_{T} (T)$,
with $x_{T}$ going away to infinity when $T\to 2\pi \omega^{-1}$,
which is the period of the linearized system at 0.

\begin{table}
\caption{This is the example table taken out of {\it The
\TeX{}book,} p.\,246}
\begin{center}
\begin{tabular}{r@{\quad}rl}
\hline
\multicolumn{1}{l}{\rule{0pt}{12pt}
                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
\hline\rule{0pt}{12pt}
8000 B.C.  &     5,000,000& \\
  50 A.D.  &   200,000,000& \\
1650 A.D.  &   500,000,000& \\
1945 A.D.  & 2,300,000,000& \\
1980 A.D.  & 4,400,000,000& \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}
%
\begin{theorem} [Ghoussoub-Preiss]\label{ghou:pre}
Assume $H(t,x)$ is
$(0,\varepsilon )$-subquadratic at
infinity for all $\varepsilon > 0$, and $T$-periodic in $t$
\begin{equation}
  H (t,\cdot )\ \ \ \ \ {\rm is\ convex}\ \ \forall t
\end{equation}
\begin{equation}
  H (\cdot ,x)\ \ \ \ \ {\rm is}\ \ T{\rm -periodic}\ \ \forall x
\end{equation}
\begin{equation}
  H (t,x)\ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n (s)s^{-1}\to \infty\ \ {\rm as}\ \ s\to \infty
\end{equation}
\begin{equation}
  \forall \varepsilon > 0\ ,\ \ \ \exists c\ :\
  H(t,x) \le \frac{\varepsilon}{2}\left\|x\right\|^{2} + c\ .
\end{equation}

Assume also that $H$ is $C^{2}$, and $H'' (t,x)$ is positive definite
everywhere. Then there is a sequence $x_{k}$, $k\in \bbbn$, of
$kT$-periodic solutions of the system
\begin{equation}
  \dot{x} = JH' (t,x)
\end{equation}
such that, for every $k\in \bbbn$, there is some $p_{o}\in\bbbn$ with:
\begin{equation}
  p\ge p_{o}\Rightarrow x_{pk} \ne x_{k}\ .
\end{equation}
\qed
\end{theorem}
%
\begin{example} [{{\rm External forcing}}]
Consider the system:
\begin{equation}
  \dot{x} = JH' (x) + f(t)
\end{equation}
where the Hamiltonian $H$ is
$\left(0,b_{\infty}\right)$-subquadratic, and the
forcing term is a distribution on the circle:
\begin{equation}
  f = \frac{d}{dt} F + f_{o}\ \ \ \ \
  {\rm with}\ \ F\in L^{2} \left(\bbbr / T\bbbz; \bbbr^{2n}\right)\ ,
\end{equation}
where $f_{o} : = T^{-1}\int_{o}^{T} f (t) dt$. For instance,
\begin{equation}
  f (t) = \sum_{k\in \bbbn} \delta_{k} \xi\ ,
\end{equation}
where $\delta_{k}$ is the Dirac mass at $t= k$ and
$\xi \in \bbbr^{2n}$ is a
constant, fits the prescription. This means that the system
$\dot{x} = JH' (x)$ is being excited by a
series of identical shocks at interval $T$.
\end{example}
%
\begin{definition}
Let $A_{\infty} (t)$ and $B_{\infty} (t)$ be symmetric
operators in $\bbbr^{2n}$, depending continuously on
$t\in [0,T]$, such that
$A_{\infty} (t) \le B_{\infty} (t)$ for all $t$.

A Borelian function
$H: [0,T]\times \bbbr^{2n} \to \bbbr$
is called
$\left(A_{\infty} ,B_{\infty}\right)$-{\it subquadratic at infinity}
if there exists a function $N(t,x)$ such that:
\begin{equation}
  H (t,x) = \frac{1}{2} \left(A_{\infty} (t) x,x\right) + N(t,x)
\end{equation}
\begin{equation}
  \forall t\ ,\ \ \ N(t,x)\ \ \ \ \
  {\rm is\ convex\ with\  respect\  to}\ \ x
\end{equation}
\begin{equation}
  N(t,x) \ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n(s)s^{-1}\to +\infty\ \ {\rm as}\ \ s\to +\infty
\end{equation}
\begin{equation}
  \exists c\in \bbbr\ :\ \ \ H (t,x) \le
  \frac{1}{2} \left(B_{\infty} (t) x,x\right) + c\ \ \ \forall x\ .
\end{equation}

If $A_{\infty} (t) = a_{\infty} I$ and
$B_{\infty} (t) = b_{\infty} I$, with
$a_{\infty} \le b_{\infty} \in \bbbr$,
we shall say that $H$ is
$\left(a_{\infty},b_{\infty}\right)$-subquadratic
at infinity. As an example, the function
$\left\|x\right\|^{\alpha}$, with
$1\le \alpha < 2$, is $(0,\varepsilon )$-subquadratic at infinity
for every $\varepsilon > 0$. Similarly, the Hamiltonian
\begin{equation}
H (t,x) = \frac{1}{2} k \left\|k\right\|^{2} +\left\|x\right\|^{\alpha}
\end{equation}
is $(k,k+\varepsilon )$-subquadratic for every $\varepsilon > 0$.
Note that, if $k<0$, it is not convex.
\end{definition}
%

\paragraph{Notes and Comments.}
The first results on subharmonics were
obtained by Rabinowitz in \cite{fo:kes:nic:tue}, who showed the existence of
infinitely many subharmonics both in the subquadratic and superquadratic
case, with suitable growth conditions on $H'$. Again the duality
approach enabled Clarke and Ekeland in \cite{may:ehr:stein} to treat the
same problem in the convex-subquadratic case, with growth conditions on
$H$ only.

Recently, Michalek and Tarantello (see \cite{fost:kes} and \cite{czaj:fitz})
have obtained lower bound on the number of subharmonics of period $kT$,
based on symmetry considerations and on pinching estimates, as in
Sect.~5.2 of this article.

%%% To ensure the bibliography has the correct style please run bibtex with the spmpsci style
%%% which is included in the Springer zip file
%%% Here we assume refs.bib would be the name of the bib file containing bibliographic info
%%% You can then copy the .bbl produced, as given in the example below
%%%
%\bibliographystyle{spmpsci}
%\bibliography{refs}
%
% ---- Bibliography ----
%
\begin{thebibliography}{6}
%

\bibitem {smit:wat}
Smith, T.F., Waterman, M.S.: Identification of common molecular subsequences.
J. Mol. Biol. 147, 195?197 (1981). \url{doi:10.1016/0022-2836(81)90087-5}

\bibitem {may:ehr:stein}
May, P., Ehrlich, H.-C., Steinke, T.: ZIB structure prediction pipeline:
composing a complex biological workflow through web services.
In: Nagel, W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006.
LNCS, vol. 4128, pp. 1148?1158. Springer, Heidelberg (2006).
\url{doi:10.1007/11823285_121}

\bibitem {fost:kes}
Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing Infrastructure.
Morgan Kaufmann, San Francisco (1999)

\bibitem {czaj:fitz}
Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid information services
for distributed resource sharing. In: 10th IEEE International Symposium
on High Performance Distributed Computing, pp. 181?184. IEEE Press, New York (2001).
\url{doi: 10.1109/HPDC.2001.945188}

\bibitem {fo:kes:nic:tue}
Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The physiology of the grid: an open grid services architecture for distributed systems integration. Technical report, Global Grid
Forum (2002)

\bibitem {onlyurl}
National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}
\end{document}

>>>>

\begin{abstract}
Many problems in  quantitative finance, uncertainty quantification, and other areas have answers of the form $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation. The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ for a well chosen sequence of nodes, $\{\bsx_0, \bsx_1, \ldots\}$.  Computing $\mu$ is equivalent to computing a $d$-dimensional integral.

@@

In mathematical terms, the outcome is $Y := f(\bsX)$, where $\bsX$ is a vector random variable of inputs determining the outcome.  The values of $f$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.  The scenarios are $\bsx_0, \bsx_1, \ldots$, which give rise to outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.  The $y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest. These scenarios are also commonly called instances of $\bsX$ or nodes.  We use these terms interchangeably.

Simple or independent and identically distributed (IID)  Monte Carlo chooses the sequence of nodes $\{\bsx_0, \bsx_1, \ldots\}$ to be IID.  Quasi-Monte Carlo (qMC) methods choose the nodes to be \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.  The difference between these two distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy (LD) sequences.

This tutorial describes what qMC is, why we would want to use qMC, and how qMC can be implemented well.  The next session provides an illustrative example of the advantages of qMC.  This is followed by a description of how LD sequences are constructed and various measures of discrepancy.  We then explain how randomization can improve qMC, how to decide what sample size, $n$, is sufficient, and how to rewrite the problem of interest in a qMC-friendly way.

@@

The aim is to approximate $\mu$ by a sample mean,
\begin{equation} \label{eq:sample_mean}
	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\bsx_i),
\end{equation}
but to do so requires some further preparation.  More on this preparation process is discussed in Section \ref{sec:reformulate}.

This integral, $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
\begin{equation}\label{eq:keisterB}
	\mu = \int_{\mathbb{R}^d} \underbrace{\pi^{d/2} \cos(\lVert \bst \rVert)}_{=:g(\bst)} \cdot \underbrace{\frac{\exp(-\lVert \bst \rVert^2)}{\pi^{d/2}}}_{\text{density of } \calN(\bszero,\mathsf{I}/2)} \, \rmd \bst = \bbE(Y) = \bbE[g(\bsT)].
\end{equation}

Since the LD sequences that underly qMC are defined to approximate the standard uniform distribution, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.  This reimagines the integral $\mu$ as the expectation of a function, $f$, of a standard uniform random variable:
\begin{multline}\label{eq:keisterC}
	\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d)\bigr) \bigr\rVert/\sqrt{2}  \Bigr)}_{=:f(\bsx)} \, \rmd \bsx \\
	= \bbE[Y]
	= \bbE[f(\bsX)], \qquad \bsX \sim \calU[0,1]^d.
\end{multline}

Consider the specific case of $d=6$ for which $\mu \approx -2.327303729298$.  We approximate $\mu$ by the sample mean defined in \eqref{eq:sample_mean} for three different kinds of nodes, $\{\bsx_i\}_{i=0}^{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}. The three kind of nodes are
\begin{enumerate}
	\renewcommand{\labelenumi}{\roman{enumi}.}
	\item Cartesian grids, $\{1/(2m), \ldots, (2m-1)/(2m) \}^6$, for $m = 2, 3, \ldots$ (blue),
	\item IID simple random  with arbitrary $n$ (orange), and
	\item Randomized LD (Sobol') sequences with $n = 1, 2, \ldots, 2^m, \ldots $ (green).
\end{enumerate}

\begin{figure}
	\centering
	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.  Grids have the largest error and LD nodes have the smallest error. \label{fig:keister-err}}
\end{figure}

Note the following from this example:
\begin{itemize}
	\item This integral is not particularly easy to evaluate numerically.  Even the best choice of nodes requires at least $n=100$ to get a relative error below $10\%$.

	\item \emph{Grid nodes.} Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.  Figure \ref{fig:grid} compares a $64$ point grid with $d = 2$ and $6$.  For $d = 6$, the possible sample sizes, $n$, are quite sparse, as shown in Figure \ref{fig:keister-err}.

@@

	\item \emph{IID nodes.}  Simple MC (orange) is a substantial improvement over grid nodes. As a reminder, for IID nodes the root mean squared error is
	\begin{equation}\label{eq:IIDerror}
		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsx))}{n}} = \frac{\mathrm{Std}(f(\bsx))}{n^{1/2}},
	\end{equation}
	where $\var$ denotes the variance and $\mathrm{Std}$ the standard deviation.  This $\mathcal{O}(n^{-1/2})$ decay is observed in Figure \ref{fig:keister-err}. For simple MC the sample size, $n$, can be any positive integer without affecting the rate of decay of the error.

	Whereas grid points collapse on top of one another when viewed in low dimensional projections, are IID points may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.  The disadvantage of IID points is that they form clusters and leave gaps.  This is because the position of any one node is independent of the position of the others.

	\item \emph{LD nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.  Typically, LD sequences have preferred sample sizes.  For this Sobol' sequence, $n$ is chosen to be positive integer powers of $2$, which tends to give better accuracy than arbitrary $n$.

	Figure \ref{fig:ld} shows typical two-dimensional projections of a $64$ LD node set.  Visually, these nodes fill the unit cube better than IID nodes.  A quantitative measure of this, the discrepancy, is defined in Section \ref{sec:discrepancy}.  Constructions of LD node set are explained in the next section.

@@

\begin{figure}
	\centering
	\includegraphics[width = 0.48\textwidth]{\figpath/64gridpts_d2.pdf}\quad
	%\includegraphics[width = 5cm]{\figpath/64gridpts_d3.pdf}\quad
	\includegraphics[width =  0.48\textwidth]{\figpath/64gridpts_d6.pdf}
	\caption{Although a two-dimensional grid covers the unit square rather well, a $d$-dimensional grid for modest $d$ does \emph{not} cover the unit cube well.  For example, one can only see four distinct points in a two dimensional projection of a six-dimensional grid with $64$ points. \label{fig:grid}}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width =\textwidth]{\figpath/64iidpts_d6.pdf}
	\caption{IID points cover the unit cube better than grid points, although one does observe clusters and gaps.  In any  projection there is  a similar looking distribution of all $64$ points \label{fig:iid}}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width =\textwidth]{\figpath/64sobolpts_d6.pdf}
	\caption{LD points cover the unit cube better even better than IID points or grids.  In any  projection there is  a similar looking distribution of all $64$ points. \label{fig:ld}}
\end{figure}

\section{LD Sequence Constructions} \label{sec:construct}
This section introduces some of the most popular LD sequences.  What sets these apart from IID sequences is that the nodes are deliberately chosen and highly correlated to one another.

\subsection{Lattice Sequences} \label{sec:lattice}
The simplest LD construction is the family of good lattice points \cite{DicEtal22a,SloJoe94}.  As a finite node set, they are defined as follows:
\begin{equation}
	\bsx_i = i\bsh/n \pmod \bsone, \qquad i = 0, \ldots, n-1
\end{equation}
where $\bsh$ is a well chosen $d$-dimensional integer vector.  Note that the set $\{\bsx_0, \ldots, \bsx_{n-1}\}$ is closed under addition modulo $\bsone$ so that it forms a group.  Figure \ref{fig:latticeconstruct} illustrates this construction for $n = 16$ and $\bsh = (1,11)$.

\begin{figure}
	\centering
	\includegraphics[width = \textwidth]{\figpath/16_lattice_construct_d2.pdf}
	\caption{The construction of a good lattice point set     \label{fig:latticeconstruct}}
\end{figure}

One disadvantage of this construction is that it is not readily extensible. For the example in Figure \ref{fig:latticeconstruct} the first $8$ points do not fill the unit square.  However, the set of points defined by even $i$ do a reasonable job.  This suggests a method for defining lattice sequences that was proposed independently by \cite{Mai81a} and \cite{HicEtal00}.

Let's start with the one-dimensional extensible lattice that is defined by the van der Corput sequence, $\phi_b(0), \phi_b(1), \ldots$.  This involves reflecting the digits of the integer $i$ in base $b$ about the decimal point.  For example, $i = 6 = 110_{2}$ give $x_6 = {}_20.011 = 3/8$, or in general
\begin{equation}
	i = i_0 + i_1b + i_2 b^2 + \cdots \implies x_i = \phi_b(i) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1).
\end{equation}

\section{Discrepancy} \label{sec:discrepancy}
\section{Randomization} \label{sec:random}
\section{Stopping Criteria} \label{sec:stop}
\section{Reformulating Our Problem} \label{sec:reformulate}
\section{Ongoing Research} \label{sec:onging}
\section{Conclusion} \label{sec:conclusion}

@@

\end{document}
